AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex

\paragraph{Results interpretation.} 
Each of the outlier ranking algorithms returns an array containing the sample
index, outlier score, and selection index (index of each sample after sorting
by outlier score). DORA provides organization and visualization modules 
intended to help users interpret and make decisions based on these outputs. 
The simplest 
 module saves a CSV of the samples sorted by their outlier score 
 (i.e., selection order). Clustering the top
  $N$ outlier selections can enable users to investigate the different types of 
 outliers that might be present in the dataset; this could be especially useful
 for separating outliers caused by noise or data artifacts vs. scientifically 
 interesting samples. We implemented the K-means and self-organizing maps 
 (SOMs) algorithms for clustering the top-N outliers. For use cases in which an
 evaluation dataset containing known outliers is available, we provide a module
 to assess how well algorithm selections correlate with known outliers. This is
 done by plotting the number of true outliers vs. number of selections made. 
We provide a module for plotting histograms of outlier scores to visualize the
 distribution of scores in the dataset (which may be, e.g., multimodal or 
 long-tailed).
In addition, we are developing a web interface which will be packaged as an
executable desktop app to further facilitate
viewing and probing of the results. Figure \ref{fig:website} shows an example 
view of the rankings across algorithms for the planetary dataset.

>>>>

\paragraph{Results interpretation.}
Each of the outlier ranking algorithms returns an array containing the sample
index, outlier score, and selection index (index of each sample after sorting
by outlier score). DORA provides organization and visualization modules
intended to help users interpret and make decisions based on these outputs.
The simplest
 module saves a CSV of the samples sorted by their outlier score 
 (i.e., selection order). Clustering the top
  $N$ outlier selections can enable users to investigate the different types of 
 outliers that might be present in the dataset; this could be especially useful
 for separating outliers caused by noise or data artifacts vs. scientifically 
 interesting samples. We implemented the K-means and self-organizing maps 
 (SOMs) algorithms for clustering the top-N outliers. For use cases in which an
 evaluation dataset containing known outliers is available, we provide a module
 to assess how well algorithm selections correlate with known outliers. This is
 done by plotting the number of true outliers vs. number of selections made. 
We provide a module for plotting histograms of outlier scores to visualize the
 distribution of scores in the dataset (which may be, e.g., multimodal or 
 long-tailed).

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{figures/doravis.png}
  \caption{A screenshot of the DORA visualizer displaying results from the
  planetary science dataset.}
  \label{fig:doravis}
\end{figure}

We developed a desktop application to easily visualize DORA results with the
Electron application framework and React frontend library. The application loads
the DORA configuration file to locate the dataset and result CSVs. Then,
it displays the ranked samples and their scores in a table sorted by their
selection order. This allows for fast and easy comparison of the results of
different methods. Figure \ref{fig:doravis} shows a screenshot of the
``Aggregate Table'' view, which displays all results from different algorithms
side-by-side.
